---
title: "figure_stats_codes"
author: "Songjun & Laura"
date: "2025-06-10"
output: html_document
---

```{r load package}
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(reshape2)
library(ggrepel)
library(vegan)
library(viridis)
library(cowplot)
library(phyloseq)
library(ANCOMBC)
library(data.table)
library(ggpubr)
library(naniar)
library(SRS)
library(ggpicrust2)
library(phyloseq)
library(ggpubr)

library(rstatix)

#many different colours
library(RColorBrewer)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))



```


#load bacteria data

```{r load data}
PMA_BA_meta_data<- read.csv("bacteria/Bacteria_meta_data_long_format_for_R_updated.csv", sep=",")

PMA_BA_feature_table <- read_tsv("bacteria/ASV_table.tsv")


PMA_BA_taxonomy <- read_tsv("bacteria/ASV_tax.gtdb.tsv")
#replace NA to 'unidentified'
PMA_BA_taxonomy[is.na(PMA_BA_taxonomy)]<-"unidentified"

PMA_BA_meta_data$ID <- as.character(PMA_BA_meta_data$ID)
#remove two NA sites B22-24
PMA_BA_meta_data<-PMA_BA_meta_data[-23,]
PMA_BA_meta_data<-PMA_BA_meta_data[-23,]
PMA_BA_meta_data$depth<-factor(PMA_BA_meta_data$depth,levels = c("22-24cm","15-17cm","8-10cm","5-6cm","1-2cm","0-1cm"))
colnames(PMA_BA_meta_data)[8]<-"Treatment"
 

PMA_BA_counts <- read.table("bacteria//ASV_table.tsv", stringsAsFactors = F, header =T)%>%

          gather(sample,count, 2:ncol(.))%>%

          dplyr::rename(ID = sample)

#Get rid of unnecessBAy chBAacter, so it will be easier to combines samples later

PMA_BA_counts$ID <- gsub('\\w+_(.*)_\\w+', '\\1', PMA_BA_counts$ID)


#PMA_BA_feature_table<-as.data.frame(PMA_BA_feature_table)

PMA_BA_feature_table %>%
  pivot_longer(!ASV_ID, names_to = "ID", values_to = "count")->PMA_BA_feature_table_long

PMA_BA_feature_table_long$ID <- gsub('\\w+_(.*)_\\w+', '\\1', PMA_BA_feature_table_long$ID)

PMA_BA_feature_table_long %>%
  pivot_wider(names_from = ID, values_from = count)->PMA_BA_feature_table



```

```{r generate asvs table}
asvs_PMA_bacteria<-PMA_BA_feature_table%>%
  gather(ID, count, 2:ncol(.)) %>% 
  filter(count > 0) %>%
  group_by(ID) %>% 
  mutate(relab = count/sum(count)) %>% 
  filter(n() > 10) %>% 
  ungroup()

```

#load archaea data

```{r load data}
PMA_AR_meta_data<- read.csv("files_archaea/Archaea_meta_data_long_format_for_R_updated.csv", sep=",")


PMA_AR_feature_table <- read_tsv("files_archaea/ASV_table.tsv")


PMA_AR_taxonomy <- read_tsv("files_archaea/ASV_tax.sbdi-gtdb.tsv")
#replace NA to 'unidentified'
PMA_AR_taxonomy[is.na(PMA_AR_taxonomy)]<-"unidentified"

PMA_AR_meta_data$ID <- as.character(PMA_AR_meta_data$ID)
#remove two NA sites B22-24
PMA_AR_meta_data<-PMA_AR_meta_data[-23,]
PMA_AR_meta_data<-PMA_AR_meta_data[-23,]
PMA_AR_meta_data$depth<-factor(PMA_AR_meta_data$depth,levels = c("22-24cm","15-17cm","8-10cm","5-6cm","1-2cm","0-1cm"))
colnames(PMA_AR_meta_data)[7]<-"Treatment"


PMA_AR_counts <- read.table("files_archaea/ASV_table.tsv", stringsAsFactors = F, header =T)%>%

          gather(sample,count, 2:ncol(.))%>%

          dplyr::rename(ID = sample)

#Get rid of unnecessary character, so it will be easier to combines samples later

PMA_AR_counts$ID <- gsub('\\w+_(.*)_\\w+', '\\1', PMA_AR_counts$ID)


#PMA_AR_feature_table<-as.data.frame(PMA_AR_feature_table)

PMA_AR_feature_table %>%
  pivot_longer(!ASV_ID, names_to = "ID", values_to = "count")->PMA_AR_feature_table_long

PMA_AR_feature_table_long$ID <- gsub('\\w+_(.*)_\\w+', '\\1', PMA_AR_feature_table_long$ID)

PMA_AR_feature_table_long %>%
  pivot_wider(names_from = ID, values_from = count)->PMA_AR_feature_table



```

```{r generate asvs table}
asvs_PMA_archaea<-PMA_AR_feature_table%>%
  gather(ID, count, 2:ncol(.)) %>% 
  filter(count > 0) %>%
  group_by(ID) %>% 
  mutate(relab = count/sum(count)) %>% 
  filter(n() > 10) %>% 
  ungroup()

```


#Figure2 cell abundance

#bacteria
```{r}

bac_fig2<-  PMA_BA_meta_data%>%
    group_by(bay, depth, Treatment)%>%
    mutate(sd=sd(Gene.copy.number.per.g.sediment.dry.weight))%>%
    mutate(mean=mean(Gene.copy.number.per.g.sediment.dry.weight))%>%
    ungroup()%>%
  ggplot(aes(x=mean, y=depth, group=Treatment, fill=Treatment))+
  scale_fill_manual(values=c("#a89b8c","#aa4465"))+
      geom_bar(stat="identity",color="black", 
           position=position_dodge(), width=.9) +
        geom_errorbar(aes(y=depth, x=mean,xmin=mean-sd, xmax=mean+sd), width=.9, color="grey25", position = position_dodge()
                )+
  theme_classic()+
  #xlim(0,5e+8)+
  ggtitle("PMA Bacteria")+
    theme(axis.text.x = element_text(angle =90, vjust = 0.5, hjust=1),
        legend.position="bottom")+
    facet_wrap(~bay, nrow=2)+
       xlab("Gene copy number/g sediment dry weight") + ylab("Depth (cm)")


bac_fig2





```

#archaea
```{r}

arc_fig2<-  PMA_AR_meta_data%>%
    group_by(bay, depth, Treatment)%>%
    mutate(sd=sd(Gene.copy.number.per.g.sediment.dry.weight.archaea))%>%
    mutate(mean=mean(Gene.copy.number.per.g.sediment.dry.weight.archaea))%>%
    ungroup()%>%
  ggplot(aes(x=mean, y=depth, group=Treatment, fill=Treatment))+
  scale_fill_manual(values=c("#a89b8c","#aa4465"))+
      geom_bar(stat="identity",color="black", 
           position=position_dodge(), width=.9) +
        geom_errorbar(aes(y=depth, x=mean,xmin=mean-sd, xmax=mean+sd), width=.9, color="grey25", position = position_dodge()
                )+
  theme_classic()+
  #xlim(0,5e+8)+
  ggtitle("PMA Archaea")+
    theme(axis.text.x = element_text(angle =90, vjust = 0.5, hjust=1),
        legend.position="bottom")+
    facet_wrap(~bay, nrow=2)+
       xlab("Gene copy number/g sediment dry weight") + ylab("Depth (cm)")


arc_fig2


```

#mix pma + no pma and bacteria+archaea for mean and sd at each depth/bay
```{r}

PMA_AR_meta_selected<-PMA_AR_meta_data[,c("ID","sample","bay","site","depth","Treatment","Gene.copy.number.per.g.sediment.dry.weight.archaea")]

PMA_BA_meta_selected<-PMA_BA_meta_data[,c("ID","sample","bay","site","depth","Treatment","Gene.copy.number.per.g.sediment.dry.weight")]
colnames(PMA_BA_meta_selected)[7]<-"Gene.copy.number.per.g.sediment.dry.weight.bacteria"

PMA_meta_selected_mixed<-PMA_AR_meta_selected%>%
  inner_join(PMA_BA_meta_selected)%>%
  mutate(gene_copy_mixed=Gene.copy.number.per.g.sediment.dry.weight.archaea+Gene.copy.number.per.g.sediment.dry.weight.bacteria)


fig2_c<-  PMA_meta_selected_mixed%>%
    group_by(bay, depth)%>%
    mutate(sd=sd(gene_copy_mixed))%>%
    mutate(mean=mean(gene_copy_mixed))%>%
    ungroup()%>%
  ggplot(aes(x=mean, y=depth))+
  #scale_fill_manual(values=c("#a89b8c","#aa4465"))+
      geom_bar(stat="identity",color="black", 
           position=position_dodge(), width=.9) +
        geom_errorbar(aes(y=depth, x=mean,xmin=mean-sd, xmax=mean+sd), width=.9, color="grey25", position = position_dodge()
                )+
  theme_classic()+
  #xlim(0,5e+8)+
  ggtitle("PMA + NO PMA & bacteria +archaea")+
    theme(axis.text.x = element_text(angle =90, vjust = 0.5, hjust=1),
        legend.position="bottom")+
    facet_wrap(~bay, nrow=2)+
       xlab("Gene copy number/g sediment dry weight") + ylab("Depth (cm)")

fig2_c

```

#mix together for figure2
```{r}

PMA_figure2<- ggarrange(bac_fig2,arc_fig2,fig2_c,
          labels = c("A", "B","C"), 
  font.label = list(size = 18),
  ncol  = 3,
  common.legend = TRUE,
  legend = "bottom")



```

#statistics cell numbers
```{r}

library(rstatix)

#data normal distributed?
shapiro.test(PMA_BA_meta_data$Gene.copy.number.per.g.sediment.dry.weight)#not normally distributed
shapiro.test(PMA_AR_meta_data$Gene.copy.number.per.g.sediment.dry.weight.archaea)#not normally distributed

#so use Wilcox test


#test cell abundance between bays
#bacteria
PMA_BA_meta_data%>%
  wilcox_test(Gene.copy.number.per.g.sediment.dry.weight ~ bay)

#archaea
PMA_AR_meta_data%>%
  wilcox_test(Gene.copy.number.per.g.sediment.dry.weight.archaea ~ bay)


#test PMA vs no PMA
#use the paired= TRUE method so it will only compare PMA VS no PMA under same sample ()

#bacteria
PMA_BA_meta_data%>%
  group_by(bay) %>%
  wilcox_test(Gene.copy.number.per.g.sediment.dry.weight ~ Treatment, paired = TRUE)

PMA_BA_meta_data%>%
  group_by(bay) %>%
  wilcox_effsize(Gene.copy.number.per.g.sediment.dry.weight ~ Treatment, paired = TRUE)

#each depth
PMA_BA_meta_data%>%
  group_by(bay, depth) %>%
  wilcox_test(Gene.copy.number.per.g.sediment.dry.weight ~ Treatment, paired = TRUE)



#archaea
PMA_AR_meta_data%>%
  group_by(bay) %>%
  wilcox_test(Gene.copy.number.per.g.sediment.dry.weight.archaea ~ Treatment, paired = TRUE)

PMA_AR_meta_data%>%
  group_by(bay) %>%
  wilcox_effsize(Gene.copy.number.per.g.sediment.dry.weight.archaea ~ Treatment, paired = TRUE)

#each depth
PMA_AR_meta_data%>%
  group_by(bay,depth) %>%
  wilcox_test(Gene.copy.number.per.g.sediment.dry.weight.archaea ~ Treatment, paired = TRUE)



#mix all data together compare between bays.

PMA_AR_meta_selected<-PMA_AR_meta_data[,c("ID","sample","bay","site","depth","Treatment","Gene.copy.number.per.g.sediment.dry.weight.archaea")]

PMA_BA_meta_selected<-PMA_BA_meta_data[,c("ID","sample","bay","site","depth","Treatment","Gene.copy.number.per.g.sediment.dry.weight")]
colnames(PMA_BA_meta_selected)[7]<-"Gene.copy.number.per.g.sediment.dry.weight.bacteria"

PMA_meta_selected_mixed<-PMA_AR_meta_selected%>%
  inner_join(PMA_BA_meta_selected)%>%
  mutate(gene_copy_mixed=Gene.copy.number.per.g.sediment.dry.weight.archaea+Gene.copy.number.per.g.sediment.dry.weight.bacteria)

PMA_meta_selected_mixed%>%
  wilcox_test(gene_copy_mixed ~ bay)

PMA_meta_selected_mixed%>%
  group_by(bay, depth)%>%
  wilcox_test(gene_copy_mixed ~ Treatment)

PMA_meta_selected_mixed%>%
  group_by(depth)%>%
  wilcox_test(gene_copy_mixed ~ bay)


```


#Figure 3 code 
```{r}

#archaea
dev_arc<-asvs_PMA_archaea[,1:3]%>% 
 inner_join(PMA_AR_meta_data, by = 'ID')%>%
  inner_join(PMA_AR_taxonomy, by = 'ASV_ID')%>%
  mutate(sample_PMA = paste(sample, Treatment, sep = "_"))%>%
  group_by(sample_PMA)%>%
  # Calculate the relative abundance of each phylum in each site
  mutate(relab = count/sum(count))%>%
  ungroup()%>% 
  filter(Kingdom == "Archaea")%>%
  mutate(relab_count= relab *Gene.copy.number.per.g.sediment.dry.weight.archaea)

dev_arc$depth <- factor(dev_arc$depth,levels=c("22-24cm","15-17cm","8-10cm","5-6cm","1-2cm","0-1cm")) 

#top20
dev_arc1 <-dev_arc %>%
  group_by(bay, depth, Treatment)%>%
  #filter(dev_relab > 0.01)%>%
  arrange(desc(relab_count))%>%
  top_n(20)%>%
  ungroup()

#rest be supplement maybe?
dev_arc2 <- dev_arc %>% anti_join(dev_arc1, by=c("ASV_ID","bay","Treatment","depth"))%>% 
  mutate(Family="zzz_Others")
dev_tot <- rbind(dev_arc1,dev_arc2)  


#modify figure table
dev_arc1_figure_table<-dev_arc1%>%
  group_by(bay, depth, Treatment, Family)%>%
  summarise(cell_ab=sum(relab_count))


dev_arc1_figure_table$Family <- gsub("ANME-1", "C.Methanophagales/ANME-1", dev_arc1_figure_table$Family)
dev_arc1_figure_table$Family <- gsub("BA1", "Bathyarchaeota/BA1", dev_arc1_figure_table$Family)
dev_arc1_figure_table$Family <- gsub("JACRNV01", "Myxococcota/JACRNV01", dev_arc1_figure_table$Family)
dev_arc1_figure_table$Family <- gsub("SG8-5", "C.Proteinplasmatales/SG8-5", dev_arc1_figure_table$Family)
dev_arc1_figure_table$Family <- gsub("SpSt-1190", "Nanoarchaeota/SpSt-1190", dev_arc1_figure_table$Family)
dev_arc1_figure_table$Family <- gsub("TCS64", "C.Bathyarchaeota/TCS64", dev_arc1_figure_table$Family)
dev_arc1_figure_table$Family <- gsub("UBA233", "C.Bathyarchaeota/UBA233", dev_arc1_figure_table$Family)
 
colnames(dev_arc1_figure_table)[4]<-"Family (Archaea)"

fig3_1 <-dev_arc1_figure_table%>% 
ggplot(aes(x = depth, y = cell_ab, fill = `Family (Archaea)`)) +
  geom_bar(stat="identity")+
   scale_fill_manual(values=c("#d34276","#d54d30","#c18942","#cfd15a","#628b44",
"#75d754","#69d4b0","#698bcd","#7b59d8","#bb75b6","#ca47c1","gray76"))+
coord_flip() +
theme(
    legend.position = 'bottom'
  )+
  xlab('depth') + ylab('gene copy numbers g/dry weight') +
   theme_classic()+
  guides(fill = guide_legend(ncol = 2)) +
  theme(
    legend.position = 'bottom',
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.text = element_text(size = 10)
  )+
   facet_wrap(~bay+Treatment)


fig3_1



#bacteria
dev_bac<-asvs_PMA_bacteria[,1:3]%>% 
 inner_join(PMA_BA_meta_data, by = 'ID')%>%
  inner_join(PMA_BA_taxonomy, by = 'ASV_ID')%>%
  mutate(sample_PMA = paste(sample, Treatment, sep = "_"))%>%
  group_by(sample_PMA)%>%
  # Calculate the relative abundance of each phylum in each site
  mutate(relab = count/sum(count))%>%
  ungroup()%>% 
  filter(Kingdom == "Bacteria")%>%
  mutate(relab_count= relab *Gene.copy.number.per.g.sediment.dry.weight)

dev_bac$depth <- factor(dev_bac$depth,levels=c("22-24cm","15-17cm","8-10cm","5-6cm","1-2cm","0-1cm")) 
str(dev_bac)

#top10
dev_bac1 <-dev_bac %>%
   group_by(bay, depth, Treatment)%>%
  #filter(dev_relab > 0.01)%>%
  arrange(desc(relab_count))%>%
  top_n(10)%>%
  ungroup()


#modify figure table
dev_bac1_figure_table<-dev_bac1%>%
  group_by(bay, depth, Treatment, Family)%>%
  summarise(cell_ab=sum(relab_count))


dev_bac1_figure_table$Family <- gsub("34-128", "Atribacterota/34-128", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("E26-bin7", "Anaerolineae/E26-bin7", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("E44-bin32", "Anaerolineae/E44-bin32", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("JAAEKA01", "Anaerolineae/JAAEKA01", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("JAFDCJ01", "Desulfobacterales/JAFDCJ01", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("JAHEKZ01", "Chromatiales/JAHEKZ01", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("RBG-16-66-30", "Aminicenantales/RBG-16-66-30", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("SM1-73", "Zixibacteria/SM1-73", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("SM23-28-2", "Chloroflexota/SM23-28-2", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("UBA11574", "Desulfobacterales/UBA11574", dev_bac1_figure_table$Family)
dev_bac1_figure_table$Family <- gsub("UBA1163", "Desulfobacterales/UBA1163", dev_bac1_figure_table$Family)


colnames(dev_bac1_figure_table)[4]<-"Family (Bacteria)"
 

fig3_2 <-dev_bac1_figure_table%>% 
ggplot(aes(x = depth, y = cell_ab, fill = `Family (Bacteria)`)) +
  geom_bar(stat="identity")+
    #scale_fill_manual(values=c("#cb0063","#fc1810","#b35300","#ffd18b","#ffc542","#413600","#b29e00","#c6d600","#daffa6","#01735d","#c8fff0","#00bfab","#002a27","#8bc1ff","#0088dc","#004fd5","#150026","#e18bff","#49005f","#e428e1","#ff2db9","#ffdced","#6f0039","gray76"))+
  scale_fill_manual(values=col_vector)+
coord_flip() +
theme(
    legend.position = 'bottom'
  )+
  xlab('depth') + ylab('gene copy numbers g/dry weight') +
   theme_classic()+
    guides(fill = guide_legend(ncol = 3)) +
  theme(
    legend.position = 'bottom',
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.text = element_text(size = 10)
  )+
   facet_wrap(~bay+Treatment)


fig3_2


PMA_figure3<-ggarrange(fig3_2, fig3_1, 
          labels = c("A", "B"), 
  font.label = list(size = 18),
  align = "hv",
  ncol  = 2,
  legend="bottom" )


```

#ANCOMBC and Figure4 code

#ANCOMBC archaea 
```{r}
asv_wide_ar<-PMA_AR_feature_table%>% 
  remove_rownames() %>% 
  column_to_rownames(var = "ASV_ID")

OTU= otu_table(asv_wide_ar, taxa_are_rows = TRUE)

ASV_tax_df<-as.data.frame(PMA_AR_taxonomy[,1:9])
rownames(ASV_tax_df)<-NULL
TAX = tax_table(as.matrix(ASV_tax_df %>% column_to_rownames(., var="ASV_ID")))
row.names(PMA_AR_meta_data)<-NULL
SAM = sample_data(PMA_AR_meta_data%>% column_to_rownames(var="ID"))

ps <- phyloseq(OTU, TAX, SAM)

ps@sam_data[["group"]] <- paste(ps@sam_data[["bay"]], ps@sam_data[["depth"]], sep="_")

#phylum level
ancombc2_PMA_archaea_phylum<-ancombc2(data = ps, tax_level = "Phylum", fix_formula = "group", group="group", rand_formula=NULL, p_adj_method = "holm", n_cl = 4, prv_cut = 0.10, struc_zero=TRUE, neg_lb=TRUE, global=TRUE, pairwise=TRUE, dunnet=TRUE)

#family level
ancombc2_PMA_archaea_family<-ancombc2(data = ps, tax_level = "Family", fix_formula = "group", group="group", rand_formula=NULL, p_adj_method = "holm", n_cl = 4, prv_cut = 0.10, struc_zero=TRUE, neg_lb=TRUE, global=TRUE, pairwise=TRUE, dunnet=TRUE)


#seperate samdata into control/warm for two dunnet results
SAM_control<-SAM[SAM$bay=="control",]
SAM_heated<-SAM[SAM$bay=="heated",]

ps_control <- phyloseq(OTU, TAX, SAM_control)
ps_heated<-phyloseq(OTU, TAX, SAM_heated)


#family level
ancombc2_PMA_archaea_family_control<-ancombc2(data = ps_control, tax_level = "Family", fix_formula = "depth", group="depth", rand_formula=NULL, p_adj_method = "holm", n_cl = 4, prv_cut = 0.10, struc_zero=TRUE, neg_lb=TRUE, global=TRUE, pairwise=TRUE, dunnet=TRUE)

ancombc2_PMA_archaea_family_heated<-ancombc2(data = ps_heated, tax_level = "Family", fix_formula = "depth", group="depth", rand_formula=NULL, p_adj_method = "holm", n_cl = 4, prv_cut = 0.10, struc_zero=TRUE, neg_lb=TRUE, global=TRUE, pairwise=TRUE, dunnet=TRUE)


PMA_archaea_family_control_dunnest<-ancombc2_PMA_archaea_family_control$res_dunn
PMA_archaea_family_heated_dunnest<-ancombc2_PMA_archaea_family_heated$res_dunn



```

#heatmap archaea
```{r}

p<- asvs_PMA_archaea%>%
  inner_join(PMA_AR_taxonomy, by="ASV_ID")%>%
  group_by(Family, ID)%>%
  summarise(relab=sum(relab))%>%
  summarise(mean_relab=sum(relab))%>%
  ungroup()%>%
  filter(!is.na(Family))%>%
  top_n(15,mean_relab)



#control archaea part

#try to add a new column top 15 taxon/family

colnames(PMA_archaea_family_control_dunnest)[1]<-"Family"

PMA_archaea_family_control_dunnest[,1:6] %>%
   left_join(
    p%>%
      transmute(Family, top_15_Family=Family),
    by="Family"
  )%>%
  replace_na(list("top_15_Family"="Other_Family"))->PMA_archaea_family_control_dunnest_wide


PMA_archaea_family_control_dunnest_wide %>%
  pivot_longer(!c(Family,top_15_Family), names_to = "group", values_to = "lfc_value")->PMA_archaea_family_control_dunnest_long

#replace group name
PMA_archaea_family_control_dunnest_long$group[PMA_archaea_family_control_dunnest_long$group == 'lfc_depth1-2cm'] <- '1-2cm'
PMA_archaea_family_control_dunnest_long$group[PMA_archaea_family_control_dunnest_long$group == 'lfc_depth5-6cm'] <- '5-6cm'
PMA_archaea_family_control_dunnest_long$group[PMA_archaea_family_control_dunnest_long$group == 'lfc_depth8-10cm'] <- '8-10cm'
PMA_archaea_family_control_dunnest_long$group[PMA_archaea_family_control_dunnest_long$group == 'lfc_depth15-17cm'] <- '15-17cm'
PMA_archaea_family_control_dunnest_long$group[PMA_archaea_family_control_dunnest_long$group == 'lfc_depth22-24cm'] <- '22-24cm'

#organize order
PMA_archaea_family_control_dunnest_long$group<- factor(PMA_archaea_family_control_dunnest_long$group,levels=c("1-2cm","5-6cm","8-10cm","15-17cm","22-24cm"))
 
 
lo = floor(min(PMA_archaea_family_control_dunnest_long$lfc_value))
up = ceiling(max(PMA_archaea_family_control_dunnest_long$lfc_value))
mid = 0#(lo + up)/2

heatmap_archaea_control_top15_family<-PMA_archaea_family_control_dunnest_long%>%
  ggplot(aes(x = top_15_Family, y = group, fill = lfc_value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white",
                       na.value = "white", midpoint = mid, limit = c(-6,6),
                       name = NULL) +
  labs(x = NULL, y = NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle("Archaea control bay surface compared to depth")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size =12))+
scale_y_discrete(limits = rev)


heatmap_archaea_control_top15_family

#ggsave("heatmap_archaea_control_top15_family.pdf",plot=heatmap_archaea_control_top15_family, width = 40, height = 20,units = "cm")



#heated part

#try to add a new column top 15 taxon/family

colnames(PMA_archaea_family_heated_dunnest)[1]<-"Family"

PMA_archaea_family_heated_dunnest[,1:6] %>%
   left_join(
    p%>%
      transmute(Family, top_15_Family=Family),
    by="Family"
  )%>%
  replace_na(list("top_15_Family"="Other_Family"))->PMA_archaea_family_heated_dunnest_wide


PMA_archaea_family_heated_dunnest_wide %>%
  pivot_longer(!c(Family,top_15_Family), names_to = "group", values_to = "lfc_value")->PMA_archaea_family_heated_dunnest_long

#replace group name
PMA_archaea_family_heated_dunnest_long$group[PMA_archaea_family_heated_dunnest_long$group == 'lfc_depth1-2cm'] <- '1-2cm'
PMA_archaea_family_heated_dunnest_long$group[PMA_archaea_family_heated_dunnest_long$group == 'lfc_depth5-6cm'] <- '5-6cm'
PMA_archaea_family_heated_dunnest_long$group[PMA_archaea_family_heated_dunnest_long$group == 'lfc_depth8-10cm'] <- '8-10cm'
PMA_archaea_family_heated_dunnest_long$group[PMA_archaea_family_heated_dunnest_long$group == 'lfc_depth15-17cm'] <- '15-17cm'
PMA_archaea_family_heated_dunnest_long$group[PMA_archaea_family_heated_dunnest_long$group == 'lfc_depth22-24cm'] <- '22-24cm'

#organize order
PMA_archaea_family_heated_dunnest_long$group<- factor(PMA_archaea_family_heated_dunnest_long$group,levels=c("1-2cm","5-6cm","8-10cm","15-17cm","22-24cm"))
 
 
lo = floor(min(PMA_archaea_family_heated_dunnest_long$lfc_value))
up = ceiling(max(PMA_archaea_family_heated_dunnest_long$lfc_value))
mid = 0#(lo + up)/2

heatmap_archaea_heated_top15_family<-PMA_archaea_family_heated_dunnest_long%>%
  ggplot(aes(x = top_15_Family, y = group, fill = lfc_value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white",
                       na.value = "white", midpoint = mid, limit = c(-6,6),
                       name = NULL) +
  labs(x = NULL, y = NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle("Archaea heated bay surface compared to depth")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size =12))+
scale_y_discrete(limits = rev)

heatmap_archaea_heated_top15_family

#ggsave("heatmap_archaea_heated_top15_family.pdf",plot=heatmap_archaea_heated_top15_family, width = 40, height = 20,units = "cm")




```

#ANCOMBC bacteria
```{r}
asv_wide_ba<-PMA_BA_feature_table%>% 
  remove_rownames() %>% 
  column_to_rownames(var = "ASV_ID")

OTU= otu_table(asv_wide_ba, taxa_are_rows = TRUE)

ASV_tax_df<-as.data.frame(PMA_BA_taxonomy[,1:9])
rownames(ASV_tax_df)<-NULL
TAX = tax_table(as.matrix(ASV_tax_df %>% column_to_rownames(., var="ASV_ID")))
row.names(PMA_BA_meta_data)<-NULL
SAM = sample_data(PMA_BA_meta_data%>% column_to_rownames(var="ID"))

ps <- phyloseq(OTU, TAX, SAM)


#seperate samdata into control/warm for two dunnet results
SAM_control<-SAM[SAM$bay=="control",]
SAM_heated<-SAM[SAM$bay=="heated",]

ps_control <- phyloseq(OTU, TAX, SAM_control)
ps_heated<-phyloseq(OTU, TAX, SAM_heated)


#family level
ancombc2_PMA_bacteria_family_control<-ancombc2(data = ps_control, tax_level = "Family", fix_formula = "depth", group="depth", rand_formula=NULL, p_adj_method = "holm", n_cl = 4, prv_cut = 0.10, struc_zero=TRUE, neg_lb=TRUE, global=TRUE, pairwise=TRUE, dunnet=TRUE)

ancombc2_PMA_bacteria_family_heated<-ancombc2(data = ps_heated, tax_level = "Family", fix_formula = "depth", group="depth", rand_formula=NULL, p_adj_method = "holm", n_cl = 4, prv_cut = 0.10, struc_zero=TRUE, neg_lb=TRUE, global=TRUE, pairwise=TRUE, dunnet=TRUE)


PMA_bacteria_family_control_dunnest<-ancombc2_PMA_bacteria_family_control$res_dunn
PMA_bacteria_family_heated_dunnest<-ancombc2_PMA_bacteria_family_heated$res_dunn

#write.csv(PMA_bacteria_family_control_dunnest, file = "~/PMA_bacteria_family_dunnest_control.csv")
#write.csv(PMA_bacteria_family_heated_dunnest, file = "~/PMA_bacteria_family_dunnest_heated.csv")



```

#heatmap bacteria
```{r}

p<- asvs_PMA_bacteria%>%
  inner_join(PMA_BA_taxonomy, by="ASV_ID")%>%
  group_by(Family, ID)%>%
  summarise(relab=sum(relab))%>%
  summarise(mean_relab=sum(relab))%>%
  ungroup()%>%
  filter(!is.na(Family))%>%
  top_n(15,mean_relab)



#control bacteria part

#try to add a new column top 15 taxon/family

colnames(PMA_bacteria_family_control_dunnest)[1]<-"Family"

PMA_bacteria_family_control_dunnest[,1:6] %>%
   left_join(
    p%>%
      transmute(Family, top_15_Family=Family),
    by="Family"
  )%>%
  replace_na(list("top_15_Family"="Other_Family"))->PMA_bacteria_family_control_dunnest_wide


PMA_bacteria_family_control_dunnest_wide %>%
  pivot_longer(!c(Family,top_15_Family), names_to = "group", values_to = "lfc_value")->PMA_bacteria_family_control_dunnest_long

#replace group name
PMA_bacteria_family_control_dunnest_long$group[PMA_bacteria_family_control_dunnest_long$group == 'lfc_depth1-2cm'] <- '1-2cm'
PMA_bacteria_family_control_dunnest_long$group[PMA_bacteria_family_control_dunnest_long$group == 'lfc_depth5-6cm'] <- '5-6cm'
PMA_bacteria_family_control_dunnest_long$group[PMA_bacteria_family_control_dunnest_long$group == 'lfc_depth8-10cm'] <- '8-10cm'
PMA_bacteria_family_control_dunnest_long$group[PMA_bacteria_family_control_dunnest_long$group == 'lfc_depth15-17cm'] <- '15-17cm'
PMA_bacteria_family_control_dunnest_long$group[PMA_bacteria_family_control_dunnest_long$group == 'lfc_depth22-24cm'] <- '22-24cm'

#organize order
PMA_bacteria_family_control_dunnest_long$group<- factor(PMA_bacteria_family_control_dunnest_long$group,levels=c("1-2cm","5-6cm","8-10cm","15-17cm","22-24cm"))
 
 
lo = floor(min(PMA_bacteria_family_control_dunnest_long$lfc_value))
up = ceiling(max(PMA_bacteria_family_control_dunnest_long$lfc_value))
mid = 0#(lo + up)/2

heatmap_bacteria_control_top15_family<-PMA_bacteria_family_control_dunnest_long%>%
  ggplot(aes(x = top_15_Family, y = group, fill = lfc_value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white",
                       na.value = "white", midpoint = mid, limit = c(-6,6),
                       name = NULL) +
  labs(x = NULL, y = NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle("Bacteria control bay surface compared to depth")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size =12))+
scale_y_discrete(limits = rev)


heatmap_bacteria_control_top15_family

#ggsave("heatmap_bacteria_control_top15_family.pdf",plot=heatmap_bacteria_control_top15_family, width = 40, height = 20,units = "cm")



#heated part

#try to add a new column top 15 taxon/family

colnames(PMA_bacteria_family_heated_dunnest)[1]<-"Family"

PMA_bacteria_family_heated_dunnest[,1:6] %>%
   left_join(
    p%>%
      transmute(Family, top_15_Family=Family),
    by="Family"
  )%>%
  replace_na(list("top_15_Family"="Other_Family"))->PMA_bacteria_family_heated_dunnest_wide


PMA_bacteria_family_heated_dunnest_wide %>%
  pivot_longer(!c(Family,top_15_Family), names_to = "group", values_to = "lfc_value")->PMA_bacteria_family_heated_dunnest_long

#replace group name
PMA_bacteria_family_heated_dunnest_long$group[PMA_bacteria_family_heated_dunnest_long$group == 'lfc_depth1-2cm'] <- '1-2cm'
PMA_bacteria_family_heated_dunnest_long$group[PMA_bacteria_family_heated_dunnest_long$group == 'lfc_depth5-6cm'] <- '5-6cm'
PMA_bacteria_family_heated_dunnest_long$group[PMA_bacteria_family_heated_dunnest_long$group == 'lfc_depth8-10cm'] <- '8-10cm'
PMA_bacteria_family_heated_dunnest_long$group[PMA_bacteria_family_heated_dunnest_long$group == 'lfc_depth15-17cm'] <- '15-17cm'
PMA_bacteria_family_heated_dunnest_long$group[PMA_bacteria_family_heated_dunnest_long$group == 'lfc_depth22-24cm'] <- '22-24cm'

#organize order
PMA_bacteria_family_heated_dunnest_long$group<- factor(PMA_bacteria_family_heated_dunnest_long$group,levels=c("1-2cm","5-6cm","8-10cm","15-17cm","22-24cm"))
 
 
lo = floor(min(PMA_bacteria_family_heated_dunnest_long$lfc_value))
up = ceiling(max(PMA_bacteria_family_heated_dunnest_long$lfc_value))
mid = 0#(lo + up)/2

heatmap_bacteria_heated_top15_family<-PMA_bacteria_family_heated_dunnest_long%>%
  ggplot(aes(x = top_15_Family, y = group, fill = lfc_value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white",
                       na.value = "white", midpoint = mid, limit = c(-6,6),
                       name = NULL) +
  labs(x = NULL, y = NULL)+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle("Bacteria heated bay surface compared to depth")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size =12))+
scale_y_discrete(limits = rev)

heatmap_bacteria_heated_top15_family

#ggsave("heatmap_bacteria_heated_top15_family.pdf",plot=heatmap_bacteria_heated_top15_family, width = 40, height = 20,units = "cm")




```

#mix as Figure 4
```{r}


PMA_heatmap_mix<-ggarrange(heatmap_archaea_heated_top15_family,
          heatmap_bacteria_heated_top15_family,
          heatmap_archaea_control_top15_family,
          heatmap_bacteria_control_top15_family,
                    labels = c("A", "B","C","D"),
                    font.label = list(size = 18),
                    common.legend = TRUE, legend = "right",
                    ncol=2, nrow=2)
PMA_heatmap_mix

PMA_heatmap_mix <- annotate_figure(
  PMA_heatmap_mix,
  left = text_grob("Depth interval", rot = 90, vjust = 1, size = 14)
)

PMA_heatmap_mix

```

#picrust analysis & Figure 5

```{r Differential abundance analysis - Heated bay - Archaea - ANCOMBC2}

#Make phyloseq object, read more: https://codeocean.com/capsule/0588632/tree/v1

meta_ancom <- metadata.heat %>% dplyr::select(sample_ID, bay, depth)%>%filter(sample_ID %in% c("P29893_1001_S1", "P29893_1002_S2", "P29893_1003_S3", "P29893_1004_S4", "P29893_1005_S5", "P29893_1006_S6", "P29893_1007_S7", "P29893_1008_S8", "P29893_1009_S9", "P29893_1010_S10", "P29893_1011_S11", "P29893_1012_S12", "P29893_1013_S13", "P29893_1014_S14", "P29893_1015_S15", "P29893_1016_S16", "P29893_1017_S17", "P29893_1018_S18", "P29893_1019_S19","P29893_1020_S20" ,"P29893_1021_S21","P29893_1022_S22", "P29893_1023_S23", "P29893_1024_S24", "P29893_1025_S25", "P29893_1026_S26", "P29893_1027_S27", "P29893_1028_S28", "P29893_1029_S29", "P29893_1030_S30", "P29893_1031_S31", "P29893_1032_S32", "P29893_1033_S33", "P29893_1034_S34", "P29893_1083_S83", "P29893_1084_S84", "P29893_1085_S85", "P29893_1086_S86", "P29893_1087_S87", "P29893_1088_S88", "P29893_1089_S89", "P29893_1090_S90", "P29893_1091_S91", "P29893_1092_S92", "P29893_1093_S93", "P29893_1094_S94"))%>% remove_rownames() %>% 
  column_to_rownames(var = "sample_ID")

# ANCOM-BC2 (SS Filter)
OTU=otu_table(strat_table_arch, taxa_are_rows = TRUE)
META = sample_data(meta_ancom)
phyl_obj <- phyloseq(OTU, META)

sample_data(phyl_obj)$depth <- as.factor(sample_data(phyl_obj)$depth)
sample_data(phyl_obj)$depth<- relevel(sample_data(phyl_obj)$depth, "0-1cm")

output = ancombc2(data = phyl_obj, assay_name = "relab",
                  fix_formula = "depth", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, lib_cut = 1000, s0_perc = 0.05,
                  group = "depth", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, 
                  dunnet = TRUE, trend = FALSE,
                  iter_control = list(tol = 1e-5, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(), 
                  mdfdr_control =  list(fwer_ctrl_method = "holm", B = 100), 
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2),
                                       solver = "ECOS",
                                       B = 10))

res_table <- as.data.frame(output$res)
#write.csv(res_table, "ANCOMBC2_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_pair = as.data.frame(output$res_pair)
#write.csv(res_pair, "ANCOMBC2_res_pair_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_global = as.data.frame(output$res_global)
#write.csv(res_global, "ANCOMBC2_res_global_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_dunn = as.data.frame(output$res_dunn)
write.csv(res_dunn, "ANCOMBC2_res_dunn_output_heated_bay_Archaea_.csv")

```

```{r ANCOM-BC2 multiple pairwise comparisons against a pre-specified group (Dunnett’s type of test) - Heated bay - Archaea - ANCOMBC2}
res_dunn <- read.csv("ANCOMBC2_res_dunn_output_heated_bay_Archaea_.csv",stringsAsFactors = FALSE, header=TRUE)


#Figure for DUNN - multiple comparison
df_fig_dunn = as.data.frame(res_dunn) %>%
    dplyr::filter(`diff_depth1.2cm`== 1 | 
                    `diff_depth5.6cm` == 1 |
                   `diff_depth8.10cm` == 1 | 
                    `diff_depth15.17cm` == 1|
                    `diff_depth22.24cm` == 1) %>%
    mutate(`lfc_depth1-2cm` = ifelse(`diff_depth1.2cm` == 1, 
                             `lfc_depth1.2cm`, 0),
           `lfc_depth5-6cm` = ifelse(`diff_depth5.6cm` == 1, 
                                   `lfc_depth5.6cm`, 0),
           `lfc_depth8-10cm` = ifelse(`diff_depth8.10cm` == 1, 
                                   `lfc_depth8.10cm`, 0),
           `lfc_depth15-17cm` = ifelse(`diff_depth15.17cm` == 1, 
                                   `lfc_depth15.17cm`, 0),
           `lfc_depth22-24cm` = ifelse(`diff_depth22.24cm` == 1, 
                                   `lfc_depth22.24cm`, 0)) %>%
    transmute(taxon, 
              `1-2cm vs. 0-1cm` = round(`lfc_depth1-2cm`, 2), 
              `5-6cm vs. 0-1cm` = round(`lfc_depth5-6cm`, 2),
              `8-10cm vs. 0-1cm` = round(`lfc_depth8-10cm`, 2),
              `15-17cm vs. 0-1cm` = round(`lfc_depth15-17cm`, 2),
              `22-24cm vs. 0-1cm` = round(`lfc_depth22-24cm`, 2) )%>%
    pivot_longer(cols = `1-2cm vs. 0-1cm`:`5-6cm vs. 0-1cm`:`8-10cm vs. 0-1cm`:`15-17cm vs. 0-1cm`:`22-24cm vs. 0-1cm`, 
                 names_to = "group", values_to = "value") %>%
    arrange(taxon)

df_fig_dunn$group<- factor(df_fig_dunn$group,levels=c("1-2cm vs. 0-1cm",
                    "5-6cm vs. 0-1cm" ,
                   "8-10cm vs. 0-1cm",
                    "15-17cm vs. 0-1cm",
                   "22-24cm vs. 0-1cm"))
  
#Filter 0 Values
#df_fig_dunn <-df_fig_dunn %>% filter(!value == "0")

#Rename taxon to KO
df_fig_dunn <- df_fig_dunn%>%dplyr::rename(KO = taxon) 

#Upload KO groups

KO_tab<-  read.delim("../Master_student_shahinez/ko00001_ed.txt")
df_fig_dunn <- df_fig_dunn %>% inner_join( KO_tab, by="KO", relationship = "many-to-many")
df_fig_dunn$gene= str_split_fixed(df_fig_dunn$gene_info,";",2)[,1]

df_fig_dunn_sub <- df_fig_dunn %>% filter(Highest %in% "09100 Metabolism") %>% filter(Secondhighest %in% "09102 Energy metabolism")%>%filter(!Thirdhighest %in% c("00190 Oxidative phosphorylation [PATH:ko00190]", "00710 Carbon fixation in photosynthetic organisms [PATH:ko00710]", "00720 Carbon fixation pathways in prokaryotes [PATH:ko00720]","00195 Photosynthesis [PATH:ko00195]","00196 Photosynthesis - antenna proteins [PATH:ko00196]" ))

df_fig_dunn_sub$group<- factor(df_fig_dunn_sub$group,levels=c("22-24cm vs. 0-1cm", "15-17cm vs. 0-1cm", "8-10cm vs. 0-1cm","5-6cm vs. 0-1cm" ,
"1-2cm vs. 0-1cm"))

df_fig_dunn_sub$gene <- factor(df_fig_dunn_sub$gene, levels=c("acs","apgM", "cofC","cofD","cofE","cofH","fbiC","fwdA, fmdA","fwdB, fmdB","fwdD, fmdD","fwdF, fmdF", "fmdE","glyA", "K01622" ,"K16306", 
"mtrA","mtrH","mttC","pmoA-amoA","pmoB-amoB","pmoC-amoC", "serA" ,"serB" ,"cdhB", "cdhC","cdhD, acsD","cdhE, acsC", "fae-hps" ,"fdoG, fdhF, fdwA","hps-phi" ,"mer" ,"mfnA, adc" ,
"mvhA, vhuA, vhcA", "mvhG, vhuG, vhcG" ,"ppc" ,"mdh" ,

"cysC", "cysH", "cysK", "cysQ","sat" ,"sir" ,"sqr" ,

"nirK",

"eno","glnA" ,"gpmB" ,"PGAM, gpmA","sseA" ,"fbaB" ,"ftr","hdrD" ,"metB", "metX" ))

lo = floor(min(df_fig_dunn_sub$value))
up = ceiling(max(df_fig_dunn_sub$value))
mid = 0#(lo + up)/2
fig_dunn_heat_arch = df_fig_dunn_sub%>%
  ggplot(aes(x = gene ,y =group, fill = value)) + 
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white", 
                       na.value = "white", midpoint = mid, limits = c(-6, 6),
                       name = NULL )+
 # geom_text(aes(gene, group, label = value), color = "black", size = 4) +
  labs(x = NULL, y = NULL, title = "Log fold changes as compared to control") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle(" Archaea heated bay surface compared to depth")+
  guides(x =  guide_axis(angle = 90))
fig_dunn_heat_arch
```

```{r control - Archaea - Picrust - ANCOMBC2}
#Subset heated bay

metadata.con <-  read.csv("amplicon.seq/archaea/Archaea_meta_data_long_format_for_R_updated_excel.csv",stringsAsFactors = FALSE, header=TRUE)%>% 
 # filter(!sample %in% "B22-24" )%>% 
  filter(!bay %in% "heated")


strat_table_arch_con <- read.delim("amplicon.seq/archaea/picrust/picrust2_out_pipeline/KO_metagenome_out/pred_metagenome_unstrat.tsv/pred_metagenome_unstrat.tsv",stringsAsFactors = FALSE, header=TRUE)%>%dplyr::select(
  -P29893_1001_S1, -P29893_1002_S2, -P29893_1003_S3, -P29893_1004_S4, -P29893_1005_S5, -P29893_1006_S6, -P29893_1007_S7, -P29893_1008_S8, -P29893_1009_S9, -P29893_1010_S10, -P29893_1011_S11, -P29893_1012_S12, -P29893_1013_S13, -P29893_1014_S14, -P29893_1015_S15, -P29893_1016_S16, -P29893_1017_S17, -P29893_1018_S18, -P29893_1019_S19,-P29893_1020_S20 ,-P29893_1021_S21,-P29893_1022_S22, -P29893_1023_S23, -P29893_1024_S24, -P29893_1025_S25, -P29893_1026_S26, -P29893_1027_S27, -P29893_1028_S28, -P29893_1029_S29, -P29893_1030_S30, -P29893_1031_S31, -P29893_1032_S32, -P29893_1033_S33, -P29893_1034_S34, -P29893_1083_S83, -P29893_1084_S84, -P29893_1085_S85, -P29893_1086_S86, -P29893_1087_S87, -P29893_1088_S88, -P29893_1089_S89, -P29893_1090_S90, -P29893_1091_S91, -P29893_1092_S92, -P29893_1093_S93, -P29893_1094_S94)%>% remove_rownames() %>% 
  column_to_rownames(var = "function.")


```

```{r DA analysis - Control bay - Archaea - ANCOMBC2}
#Make phyloseq object, read more: https://codeocean.com/capsule/0588632/tree/v1

meta_ancom <- metadata.con %>%  remove_rownames() %>% 
  column_to_rownames(var = "sample_ID")

# ANCOM-BC2 (SS Filter)
OTU=otu_table(strat_table_arch_con, taxa_are_rows = TRUE)
META = sample_data(meta_ancom)
phyl_obj <- phyloseq(OTU, META)

sample_data(phyl_obj)$depth <- as.factor(sample_data(phyl_obj)$depth)
sample_data(phyl_obj)$depth<- relevel(sample_data(phyl_obj)$depth, "0-1cm")

output = ancombc2(data = phyl_obj, assay_name = "relab",
                  fix_formula = "depth", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, lib_cut = 1000, s0_perc = 0.05,
                  group = "depth", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, 
                  dunnet = TRUE, trend = FALSE,
                  iter_control = list(tol = 1e-5, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(), 
                  mdfdr_control =  list(fwer_ctrl_method = "holm", B = 100), 
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2),
                                       solver = "ECOS",
                                       B = 10))

res_table <- as.data.frame(output$res)
#write.csv(res_table, "ANCOMBC2_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_pair = as.data.frame(output$res_pair)
#write.csv(res_pair, "ANCOMBC2_res_pair_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_global = as.data.frame(output$res_global)
#write.csv(res_global, "ANCOMBC2_res_global_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_dunn = as.data.frame(output$res_dunn)
write.csv(res_dunn, "ANCOMBC2_res_dunn_output_controlbay_Archaea_.csv")

```

```{r ANCOMBC2 DUnnets Test - Control Bay- Archaea- ANCOMBC2}
res_dunn <- read.csv("ANCOMBC2_res_dunn_output_controlbay_Archaea_.csv",stringsAsFactors = FALSE, header=TRUE)





#Figure for DUNN - multiple comparison
df_fig_dunn = as.data.frame(res_dunn) %>%
    dplyr::filter(`diff_depth1.2cm`== 1 | 
                    `diff_depth5.6cm` == 1 |
                   `diff_depth8.10cm` == 1 | 
                    `diff_depth15.17cm` == 1|
                    `diff_depth22.24cm` == 1) %>%
    mutate(`lfc_depth1-2cm` = ifelse(`diff_depth1.2cm` == 1, 
                             `lfc_depth1.2cm`, 0),
           `lfc_depth5-6cm` = ifelse(`diff_depth5.6cm` == 1, 
                                   `lfc_depth5.6cm`, 0),
           `lfc_depth8-10cm` = ifelse(`diff_depth8.10cm` == 1, 
                                   `lfc_depth8.10cm`, 0),
           `lfc_depth15-17cm` = ifelse(`diff_depth15.17cm` == 1, 
                                   `lfc_depth15.17cm`, 0),
           `lfc_depth22-24cm` = ifelse(`diff_depth22.24cm` == 1, 
                                   `lfc_depth22.24cm`, 0)) %>%
    transmute(taxon, 
              `1-2cm vs. 0-1cm` = round(`lfc_depth1-2cm`, 2), 
              `5-6cm vs. 0-1cm` = round(`lfc_depth5-6cm`, 2),
              `8-10cm vs. 0-1cm` = round(`lfc_depth8-10cm`, 2),
              `15-17cm vs. 0-1cm` = round(`lfc_depth15-17cm`, 2),
              `22-24cm vs. 0-1cm` = round(`lfc_depth22-24cm`, 2) )%>%
    pivot_longer(cols = `1-2cm vs. 0-1cm`:`5-6cm vs. 0-1cm`:`8-10cm vs. 0-1cm`:`15-17cm vs. 0-1cm`:`22-24cm vs. 0-1cm`, 
                 names_to = "group", values_to = "value") %>%
    arrange(taxon)

df_fig_dunn$group<- factor(df_fig_dunn$group,levels=c("1-2cm vs. 0-1cm",
                    "5-6cm vs. 0-1cm" ,
                   "8-10cm vs. 0-1cm",
                    "15-17cm vs. 0-1cm",
                   "22-24cm vs. 0-1cm"))
  
#Filter 0 Values
#df_fig_dunn <-df_fig_dunn %>% filter(!value == "0")

#Rename taxon to KO
df_fig_dunn <- df_fig_dunn%>%dplyr::rename(KO = taxon) 

#Upload KO groups
KO_tab<-  read.delim("../Master_student_shahinez/ko00001_ed.txt")
df_fig_dunn <- df_fig_dunn %>% inner_join( KO_tab, by="KO", relationship = "many-to-many")
df_fig_dunn$gene= str_split_fixed(df_fig_dunn$gene_info,";",2)[,1]

df_fig_dunn_sub <- df_fig_dunn %>% filter(Highest %in% "09100 Metabolism") %>% filter(Secondhighest %in% "09102 Energy metabolism")%>%filter(!Thirdhighest %in% c("00190 Oxidative phosphorylation [PATH:ko00190]", "00710 Carbon fixation in photosynthetic organisms [PATH:ko00710]", "00720 Carbon fixation pathways in prokaryotes [PATH:ko00720]","00195 Photosynthesis [PATH:ko00195]","00196 Photosynthesis - antenna proteins [PATH:ko00196]" ))




df_fig_dunn_sub$group<- factor(df_fig_dunn_sub$group,levels=c("22-24cm vs. 0-1cm", "15-17cm vs. 0-1cm", "8-10cm vs. 0-1cm","5-6cm vs. 0-1cm" ,
"1-2cm vs. 0-1cm"))

df_fig_dunn_sub$gene <- factor(df_fig_dunn_sub$gene, levels=c("ACSS1_2, acs","apgM", "cofC, fbiD","cofD","cofE, fbiB","cofH","fbiC","fbiB","fwdA, fmdA","fwdB, fmdB","fwdD, fmdD","fwdF, fmdF", "fwdE, fmdE","fwdC, fmdC","glyA, SHMT", "K01622" ,"K16306", 
"mtrA","mtrH","mttC","pmoA-amoA","pmoB-amoB","pmoC-amoC", "serA, PHGDH" ,"serB, PSPH" ,"cdhB", "cdhC","cdhD, acsD","cdhE, acsC", "fae-hps" ,"fdoG, fdhF, fdwA","hps-phi" ,"mer" ,"mfnA, adc" ,"mvhA, vhuA, vhcA", "mvhG, vhuG, vhcG" ,"ppc" ,"mdh" ,

"cysC", "cysH", "cysK", "cysQ, MET22, BPNT1","sat, met3" ,"sir" ,"sqr" ,

"nirK",

"ENO, eno","glnA, GLUL" ,"gpmB" ,"PGAM, gpmA","TST, MPST, sseA" ,"fbaB" ,"ftr","hdrD" ,"metB", "metX" ))

lo = floor(min(df_fig_dunn_sub$value))
up = ceiling(max(df_fig_dunn_sub$value))
mid = 0#(lo + up)/2
fig_dunn_con_arch = df_fig_dunn_sub%>%
  ggplot(aes(x = gene ,y =group, fill = value)) + 
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white", 
                       na.value = "white", midpoint = mid, limit =  c(-6, 6),
                       name = NULL) +
# geom_text(aes(gene, group, label = value), color = "black", size = 4) +
  labs(x = NULL, y = NULL, title = "Log fold changes as compared to control") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle(" Archaea control bay surface compared to depth")+
  guides(x =  guide_axis(angle = 90))
fig_dunn_con_arch
```

#Heated bay - Bacteria
#only selected energy metabolism due to high annotation of potential functions

```{r heated bay - bacteria}
metadata.heat.bac <-  read.csv("amplicon.seq/bacteria/Bacteria_meta_data_long_format_for_R_new_excel.csv",stringsAsFactors = FALSE, header=TRUE)%>% 
 # filter(!sample %in% "B22-24" )%>% 
  filter(bay %in% "heated")

list_heat <- metadata.heat.bac %>% dplyr::select(name)

strat_table_bac.heat <- read.delim("amplicon.seq/bacteria/picrust/picrust2_out_pipeline/KO_metagenome_out/pred_metagenome_unstrat.tsv/pred_metagenome_unstrat.tsv",stringsAsFactors = FALSE, header=TRUE)%>%dplyr::select('function.',P29892_1002_S2,P29892_1001_S1,P29892_1004_S4,				P29892_1003_S3,				P29892_1006_S6,				P29892_1005_S5,				P29892_1008_S8,				P29892_1007_S7,				P29892_1010_S10,				P29892_1009_S9,P29892_1012_S12,				P29892_1011_S11,				P29892_1014_S14,				P29892_1013_S13,				P29892_1016_S16,				P29892_1015_S15,				P29892_1018_S18,				P29892_1017_S17,				P29892_1020_S20,				P29892_1019_S19,P29892_1022_S22,				P29892_1021_S21,				P29892_1024_S24,				P29892_1023_S23,				P29892_1026_S26,				P29892_1025_S25,				P29892_1028_S28,				P29892_1027_S27,P29892_1030_S30,				
P29892_1029_S29,				P29892_1032_S32,				P29892_1031_S31,				P29892_1034_S34,				P29892_1033_S33,			P29892_1084_S84,				P29892_1083_S83,				P29892_1086_S86,				P29892_1085_S85,P29892_1088_S88,				P29892_1087_S87,				P29892_1090_S90,				P29892_1089_S89,				P29892_1092_S92,				P29892_1091_S91,				P29892_1094_S94,				P29892_1093_S93)%>% remove_rownames() %>% 
  column_to_rownames(var = "function.")

```

```{r DA analysis - heated bay - Bacteria - ANCOMBC2}
#Make phyloseq object, read more: https://codeocean.com/capsule/0588632/tree/v1

meta_ancom <- metadata.heat.bac %>%  filter(!sample %in% "B22-24")%>%
remove_rownames() %>% column_to_rownames(var = "name")

# ANCOM-BC2 (SS Filter)
OTU=otu_table(strat_table_bac.heat, taxa_are_rows = TRUE)
META = sample_data(meta_ancom)
phyl_obj <- phyloseq(OTU, META)

sample_data(phyl_obj)$depth <- as.factor(sample_data(phyl_obj)$depth)
sample_data(phyl_obj)$depth<- relevel(sample_data(phyl_obj)$depth, "0-1cm")

output = ancombc2(data = phyl_obj, assay_name = "relab",
                  fix_formula = "depth", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, lib_cut = 1000, s0_perc = 0.05,
                  group = "depth", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, 
                  dunnet = TRUE, trend = FALSE,
                  iter_control = list(tol = 1e-5, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(), 
                  mdfdr_control =  list(fwer_ctrl_method = "holm", B = 100), 
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2),
                                       solver = "ECOS",
                                       B = 10))

res_table <- as.data.frame(output$res)
#write.csv(res_table, "ANCOMBC2_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_pair = as.data.frame(output$res_pair)
#write.csv(res_pair, "ANCOMBC2_res_pair_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_global = as.data.frame(output$res_global)
#write.csv(res_global, "ANCOMBC2_res_global_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_dunn = as.data.frame(output$res_dunn)
write.csv(res_dunn, "ANCOMBC2_res_dunn_output_heatedbay_bacteria_.csv")

```

```{r ANCOMBC2 DUNNETS TEST - Heated bay - Bacteria }
res_dunn <- read.csv("ANCOMBC2_res_dunn_output_heatedbay_bacteria_.csv",stringsAsFactors = FALSE, header=TRUE)



#Figure for DUNN - multiple comparison
df_fig_dunn = as.data.frame(res_dunn) %>%
    dplyr::filter(`diff_depth1.2cm`== 1 | 
                    `diff_depth5.6cm` == 1 |
                   `diff_depth8.10cm` == 1 | 
                    `diff_depth15.17cm` == 1|
                    `diff_depth22.24cm` == 1) %>%
    mutate(`lfc_depth1-2cm` = ifelse(`diff_depth1.2cm` == 1, 
                             `lfc_depth1.2cm`, 0),
           `lfc_depth5-6cm` = ifelse(`diff_depth5.6cm` == 1, 
                                   `lfc_depth5.6cm`, 0),
           `lfc_depth8-10cm` = ifelse(`diff_depth8.10cm` == 1, 
                                   `lfc_depth8.10cm`, 0),
           `lfc_depth15-17cm` = ifelse(`diff_depth15.17cm` == 1, 
                                   `lfc_depth15.17cm`, 0),
           `lfc_depth22-24cm` = ifelse(`diff_depth22.24cm` == 1, 
                                   `lfc_depth22.24cm`, 0)) %>%
    transmute(taxon, 
              `1-2cm vs. 0-1cm` = round(`lfc_depth1-2cm`, 2), 
              `5-6cm vs. 0-1cm` = round(`lfc_depth5-6cm`, 2),
              `8-10cm vs. 0-1cm` = round(`lfc_depth8-10cm`, 2),
              `15-17cm vs. 0-1cm` = round(`lfc_depth15-17cm`, 2),
              `22-24cm vs. 0-1cm` = round(`lfc_depth22-24cm`, 2) )%>%
    pivot_longer(cols = `1-2cm vs. 0-1cm`:`5-6cm vs. 0-1cm`:`8-10cm vs. 0-1cm`:`15-17cm vs. 0-1cm`:`22-24cm vs. 0-1cm`, 
                 names_to = "group", values_to = "value") %>%
    arrange(taxon)

df_fig_dunn$group<- factor(df_fig_dunn$group,levels=c("1-2cm vs. 0-1cm",
                    "5-6cm vs. 0-1cm" ,
                   "8-10cm vs. 0-1cm",
                    "15-17cm vs. 0-1cm",
                   "22-24cm vs. 0-1cm"))
  
#Filter 0 Values
#df_fig_dunn <-df_fig_dunn %>% filter(!value == "0")

#Rename taxon to KO
df_fig_dunn <- df_fig_dunn%>%dplyr::rename(KO = taxon) 

#Upload KO groups

KO_tab<-  read.delim("../Master_student_shahinez/ko00001_ed.txt")


df_fig_dunn <- df_fig_dunn %>% inner_join( KO_tab, by="KO", relationship = "many-to-many")
df_fig_dunn$gene= str_split_fixed(df_fig_dunn$gene_info,";",2)[,1]

df_fig_dunn_sub <- df_fig_dunn %>% filter(Highest %in% "09100 Metabolism") %>% filter(Secondhighest %in% "09102 Energy metabolism")%>%filter(!Thirdhighest %in% c("00190 Oxidative phosphorylation [PATH:ko00190]", "00710 Carbon fixation in photosynthetic organisms [PATH:ko00710]", "00720 Carbon fixation pathways in prokaryotes [PATH:ko00720]","00195 Photosynthesis [PATH:ko00195]","00196 Photosynthesis - antenna proteins [PATH:ko00196]" ))%>%filter(gene %in% c("hprA", "fdhA", "cdhA", "cdhC","cdhB", "mtd", "mcrA", "mcrB", "mcrG", "mtrA", "mtrB", "mtrC", "mtrD", "mtrE", "mtrF", "mtrG", "mtrH", "mcrC", "mcrD", "mtaB", "hdrD", "hdrE", "mtkB", "fdhB",
                                                                                                                                                                                                                                                                                                                                                                                                                                                                      "nasB", "nirD", "nirA", "narB", "narG, narZ, nxrA", "nasA", "narI, narV", "nosZ", "napA", "napB", "norB", 
                                                                                                                                                                                                                                                                                                                                                                                                                                                                      "sir", "dmsA", "dmsB", "phsA, psrA", "soxA", "soxX", "soxB", "soxY", "soxZ", "fccB", "fccA"))


df_fig_dunn_sub$gene <- factor(df_fig_dunn_sub$gene, levels=c("hprA","fdhA","fdhB","cdhA", "cdhC","cdhB","mtd","mcrA", "mcrB", "mcrG", "mcrC", "mcrD","mtrA", "mtrB", "mtrC", "mtrD", "mtrE", "mtrF", "mtrG", "mtrH","mtaB", "hdrD", "hdrE", "mtkB","nasB", "nirD", "nirA", "narB", "narG, narZ, nxrA", "nasA", "narI, narV", "nosZ", "napA", "napB", "norB","sir", "dmsA", "dmsB", "phsA, psrA", "soxA", "soxX", "soxB", "soxY", "soxZ", "fccB", "fccA"  ))



df_fig_dunn_sub$group<- factor(df_fig_dunn_sub$group,levels=c("22-24cm vs. 0-1cm", "15-17cm vs. 0-1cm", "8-10cm vs. 0-1cm","5-6cm vs. 0-1cm" ,
"1-2cm vs. 0-1cm"))


lo = floor(min(df_fig_dunn_sub$value))
up = ceiling(max(df_fig_dunn_sub$value))
mid = 0#(lo + up)/2
fig_dunn_heat_bac = df_fig_dunn_sub%>%
  ggplot(aes(x = gene, y = group, fill = value)) + 
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4",mid = "white", 
                   na.value = "white", midpoint = mid, limit =  c(-6, 6),
                 name = NULL) +

#  geom_text(aes(gene, group, label = value), color = "black", size = 4) +
  labs(x = NULL, y = NULL, title = "Log fold changes as compared to control") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle(" Bacteria heated bay surface compared to depth")+
  guides(x =  guide_axis(angle = 90))+
   annotate("rect", xmin = 0.5, xmax = 24.5, ymin = 0.5, ymax = 5.5,
             alpha = 0, color= "black", linetype="solid", size=0.8)+
  annotate("rect", xmin = 24.5, xmax = 35.5, ymin = 0.5, ymax = 5.5,
             alpha = 0, color= "black", linetype="solid", size=0.8)+
  annotate("rect", xmin = 35.5, xmax = 46.5, ymin = 0.5, ymax = 5.5,
             alpha = 0, color= "black", linetype="solid", size=0.8)
  
fig_dunn_heat_bac



```

```{r control bay - bacteria}
metadata.con.bac <-  read.csv("amplicon.seq/bacteria/Bacteria_meta_data_long_format_for_R_new_excel.csv",stringsAsFactors = FALSE, header=TRUE)%>% 
 # filter(!sample %in% "B22-24" )%>% 
  filter(bay %in% "control")

list_con <- metadata.con.bac %>% dplyr::select(name)

strat_table_bac.con <- read.delim("amplicon.seq/bacteria/picrust/picrust2_out_pipeline/KO_metagenome_out/pred_metagenome_unstrat.tsv/pred_metagenome_unstrat.tsv",stringsAsFactors = FALSE, header=TRUE)%>%dplyr::select('function.',	P29892_1036_S36,P29892_1035_S35,P29892_1038_S38,P29892_1037_S37,P29892_1040_S40,P29892_1039_S39,P29892_1042_S42,P29892_1041_S41,P29892_1044_S44,P29892_1043_S43,P29892_1046_S46,P29892_1045_S45,P29892_1048_S48,P29892_1047_S47,P29892_1050_S50,P29892_1049_S49,P29892_1052_S52,P29892_1051_S51,P29892_1054_S54,P29892_1053_S53,P29892_1056_S56,P29892_1055_S55,P29892_1058_S58,P29892_1057_S57,P29892_1060_S60,P29892_1059_S59,P29892_1062_S62,P29892_1061_S61,P29892_1064_S64,P29892_1063_S63,P29892_1066_S66,P29892_1065_S65,P29892_1068_S68,P29892_1067_S67,P29892_1070_S70,P29892_1069_S69,P29892_1072_S72,P29892_1071_S71,P29892_1074_S74,P29892_1073_S73,P29892_1076_S76,P29892_1075_S75,P29892_1078_S78,P29892_1077_S77,P29892_1080_S80,P29892_1079_S79,P29892_1082_S82,P29892_1081_S81)%>% remove_rownames() %>% 
  column_to_rownames(var = "function.")
```

```{r DA analysis - control bay - bacteria - ANCOMBC2}
#Make phyloseq object, read more: https://codeocean.com/capsule/0588632/tree/v1

meta_ancom <- metadata.con.bac %>%  filter(!sample %in% "B22-24")%>%
remove_rownames() %>% column_to_rownames(var = "name")

# ANCOM-BC2 (SS Filter)
OTU=otu_table(strat_table_bac.con, taxa_are_rows = TRUE)
META = sample_data(meta_ancom)
phyl_obj <- phyloseq(OTU, META)

sample_data(phyl_obj)$depth <- as.factor(sample_data(phyl_obj)$depth)
sample_data(phyl_obj)$depth<- relevel(sample_data(phyl_obj)$depth, "0-1cm")

output = ancombc2(data = phyl_obj, assay_name = "relab",
                  fix_formula = "depth", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, lib_cut = 1000, s0_perc = 0.05,
                  group = "depth", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, 
                  dunnet = TRUE, trend = FALSE,
                  iter_control = list(tol = 1e-5, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(), 
                  mdfdr_control =  list(fwer_ctrl_method = "holm", B = 100), 
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2),
                                       solver = "ECOS",
                                       B = 10))

res_table <- as.data.frame(output$res)
#write.csv(res_table, "ANCOMBC2_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_pair = as.data.frame(output$res_pair)
#write.csv(res_pair, "ANCOMBC2_res_pair_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_global = as.data.frame(output$res_global)
#write.csv(res_global, "ANCOMBC2_res_global_output_User.ID_SAR_group_Family_Taxa_Level_reference_control0.csv")

res_dunn = as.data.frame(output$res_dunn)
write.csv(res_dunn, "ANCOMBC2_res_dunn_output_controlbay_bacteria_.csv")
```

```{r ANCOMBC DUNNETS TEST CONTROL BAY BACTERIA}
res_dunn <- read.csv("ANCOMBC2_res_dunn_output_controlbay_bacteria_.csv",stringsAsFactors = FALSE, header=TRUE)


#Figure for DUNN - multiple comparison
df_fig_dunn = as.data.frame(res_dunn) %>%
    dplyr::filter(`diff_depth1.2cm`== 1 | 
                    `diff_depth5.6cm` == 1 |
                   `diff_depth8.10cm` == 1 | 
                    `diff_depth15.17cm` == 1|
                    `diff_depth22.24cm` == 1) %>%
    mutate(`lfc_depth1-2cm` = ifelse(`diff_depth1.2cm` == 1, 
                             `lfc_depth1.2cm`, 0),
           `lfc_depth5-6cm` = ifelse(`diff_depth5.6cm` == 1, 
                                   `lfc_depth5.6cm`, 0),
           `lfc_depth8-10cm` = ifelse(`diff_depth8.10cm` == 1, 
                                   `lfc_depth8.10cm`, 0),
           `lfc_depth15-17cm` = ifelse(`diff_depth15.17cm` == 1, 
                                   `lfc_depth15.17cm`, 0),
           `lfc_depth22-24cm` = ifelse(`diff_depth22.24cm` == 1, 
                                   `lfc_depth22.24cm`, 0)) %>%
    transmute(taxon, 
              `1-2cm vs. 0-1cm` = round(`lfc_depth1-2cm`, 2), 
              `5-6cm vs. 0-1cm` = round(`lfc_depth5-6cm`, 2),
              `8-10cm vs. 0-1cm` = round(`lfc_depth8-10cm`, 2),
              `15-17cm vs. 0-1cm` = round(`lfc_depth15-17cm`, 2),
              `22-24cm vs. 0-1cm` = round(`lfc_depth22-24cm`, 2) )%>%
    pivot_longer(cols = `1-2cm vs. 0-1cm`:`5-6cm vs. 0-1cm`:`8-10cm vs. 0-1cm`:`15-17cm vs. 0-1cm`:`22-24cm vs. 0-1cm`, 
                 names_to = "group", values_to = "value") %>%
    arrange(taxon)

df_fig_dunn$group<- factor(df_fig_dunn$group,levels=c("1-2cm vs. 0-1cm",
                    "5-6cm vs. 0-1cm" ,
                   "8-10cm vs. 0-1cm",
                    "15-17cm vs. 0-1cm",
                   "22-24cm vs. 0-1cm"))
  
#Filter 0 Values
#df_fig_dunn <-df_fig_dunn %>% filter(!value == "0")

#Rename taxon to KO
df_fig_dunn <- df_fig_dunn%>%dplyr::rename(KO = taxon) 

#Upload KO groups

KO_tab<-  read.delim("../Master_student_shahinez/ko00001_ed.txt")


df_fig_dunn <- df_fig_dunn %>% inner_join( KO_tab, by="KO", relationship = "many-to-many")

df_fig_dunn$gene= str_split_fixed(df_fig_dunn$gene_info,";",2)[,1]

df_fig_dunn_sub <- df_fig_dunn %>% filter(Highest %in% "09100 Metabolism") %>% filter(Secondhighest %in% "09102 Energy metabolism")%>%filter(!Thirdhighest %in% c("00190 Oxidative phosphorylation [PATH:ko00190]", "00710 Carbon fixation in photosynthetic organisms [PATH:ko00710]", "00720 Carbon fixation pathways in prokaryotes [PATH:ko00720]","00195 Photosynthesis [PATH:ko00195]","00196 Photosynthesis - antenna proteins [PATH:ko00196]" ))%>%filter(gene %in% c("hprA", "fdhA", "cdhA", "cdhC","cdhB", "mtd", "mcrA", "mcrB", "mcrG", "mtrA", "mtrB", "mtrC", "mtrD", "mtrE", "mtrF", "mtrG", "mtrH", "mcrC", "mcrD", "mtaB", "hdrD", "hdrE", "mtkB", "fdhB",
                                                                                                                                                                                                                                                                                                                                                                                                                                                                      "nasB", "nirD", "nirA", "narB", "narG, narZ, nxrA", "nasA", "narI, narV", "nosZ", "napA", "napB", "norB", 
                                                                                                                                                                                                                                                                                                                                                                                                                                                                      "sir", "dmsA", "dmsB", "phsA, psrA", "soxA", "soxX", "soxB", "soxY", "soxZ", "fccB", "fccA"))


df_fig_dunn_sub$gene <- factor(df_fig_dunn_sub$gene, levels=c("hprA","fdhA","fdhB","cdhA", "cdhC","cdhB","mtd","mcrA", "mcrB", "mcrG", "mcrC", "mcrD","mtrA", "mtrB", "mtrC", "mtrD", "mtrE", "mtrF", "mtrG", "mtrH","mtaB", "hdrD", "hdrE", "mtkB","nasB", "nirD", "nirA", "narB", "narG, narZ, nxrA", "nasA", "narI, narV", "nosZ", "napA", "napB", "norB","sir", "dmsA", "dmsB", "phsA, psrA", "soxA", "soxX", "soxB", "soxY", "soxZ", "fccB", "fccA"  ))



df_fig_dunn_sub$group<- factor(df_fig_dunn_sub$group,levels=c("22-24cm vs. 0-1cm", "15-17cm vs. 0-1cm", "8-10cm vs. 0-1cm","5-6cm vs. 0-1cm" ,
"1-2cm vs. 0-1cm"))


lo = floor(min(df_fig_dunn_sub$value))
up = ceiling(max(df_fig_dunn_sub$value))
mid = 0#(lo + up)/2
fig_dunn_con_bac = df_fig_dunn_sub%>%
  ggplot(aes(x = gene, y = group, fill = value)) + 
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "darkolivegreen3", high = "indianred4", mid = "white", 
                       na.value = "white", midpoint = mid, limit = c(-6,6),
                       name = NULL) +
 # geom_text(aes(gene, group, label = value), color = "black", size = 4) +
  labs(x = NULL, y = NULL, title = "Log fold changes as compared to control") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+ggtitle(" Bacteria control bay surface compared to depth")+
    guides(x =  guide_axis(angle = 90))+
   annotate("rect", xmin = 0.5, xmax = 21.5, ymin = 0.5, ymax = 5.5,
             alpha = 0, color= "black", linetype="solid", size=0.8)+
  annotate("rect", xmin = 21.5, xmax = 31.5, ymin = 0.5, ymax = 5.5,
             alpha = 0, color= "black", linetype="solid", size=0.8)+
  annotate("rect", xmin = 31.5, xmax = 40.5, ymin = 0.5, ymax = 5.5,
             alpha = 0, color= "black", linetype="solid", size=0.8)
fig_dunn_con_bac



```

```{r combine}

#bacteria
bac <-ggarrange(fig_dunn_heat_bac,fig_dunn_con_bac, nrow=2, labels = c("A", "B"),common.legend = TRUE , legend= "right")
arch <- ggarrange(fig_dunn_heat_arch, fig_dunn_con_arch, nrow=2, labels = c("C","D"), legend="right", common.legend = T)

ggarrange(bac, arch, ncol=2, common.legend = T, legend="right")
```

